Measuring starspot properties by forward modeling IGRINS spectra of \name

### 2016-06-20 document/ms.tex
Writing and revising the intro and methods section, move the detailed methods to the Appendix

\title{IGRINS spectra of \name}

@@

\subsection{Pre Main Sequence HR diagram spread}

For a single presumably coeval stellar cluster, pre main-sequence HR diagrams exhibit large ($\delta t_{\ast} \sim t_{\ast}$) apparent spreads in age \citep[e.g.][]{2011A&A...534A..83R}. The large spreads in these pre main-sequence HR diagrams is controversial: some take it as evidence for intrinsic age spreads, some take it as evidence for non static accretion history \citep{2009ApJ...702L..27B, 2010ARA&A..48..581S}.

@@

\subsection{Previous studies of Starspots}

Meanwhile, sunspots are probably responsible for biases in stellar effective temperatures derived by different methods.  For example, the APOGEE spectrograph ($1.5-1.70 \;\mu$m at $R=22,500$) measured effective temperatures for 3493 young stars finding offsets in $\teff$ of typically 200$-$500 K and as high as 1000 K compared to previous studies \citep{2014ApJ...794..125C}.  Low resolution optical spectroscopy shows a typical spread of 200 K in the spectral type-to-effective-temperature conversion scale \citep{2014ApJ...786...97H}.  Stellar evolution models including the effect of starspots can make a coeval 10 Myr population exhibit apparent age spreads of 3$-$10 Myr, with derived masses biased towards lower masses.  The observations of sunspots are lagging behind the theory \citep{2015ApJ...807..174S}.

The importance of starspots needs to be empirically evaluated to assess systematic biases of pre-main sequence stellar evolutionary models.

Modeling the periodic light curves of spotted stars can answer these questions: ``What evidence is there to support the existence of starspots?  What constraint can be placed jointly upon the effective temperature contrast and areal coverage of the starspots?  What is the longitudinal asymmetry of the spots?  ''.  As pointed out by CITE XX, cyclical light curve variations indicate only the longitudinally asymmetric component of the star spots.  Photometric modeling of the light curve produces degenerate estimates in the temperature contrast and areal coverage fraction.  For example, CITE XX achieved XX.

By modeling the composite \emph{spectrum} of a spotted star, it is possible to directly measure the collective areal coverage and effective temperature contrast.  A stellar spectrum with a longitudinally symmetric distribution of starspots, or a pole-on star with a single spot, or a non-rotating star would all show zero amplitude of photometric variability, yet they all contain star spots.  Spectroscopic modeling can still measure the starspot properties in these cases.

@@

\subsection{Previous studies of \name}

In this paper we characterize the starspot coverage and temperature contrast of \name using full spectrum fitting.  We chose \name for its likelihood of large areal coverage of starspots, as evinced by large-amplitude photometric variability  detected in $BVRI$ bands.  Its period is 3.37$\pm$0.01 days \citep{1993AJ....106.1608V,1994IBVS.4042....1G,grankin08,xiao12}.  Recent optical spectro-polarimetry showed evidence for hot or cool starspots covering an estimated $25\%$ of the stellar surface \citep{2014MNRAS.444.3220D}.  The source has little or no accretion ($\log{\dot M} < 8.1$), no detection of mid-infrared nor mm excess, and no stellar companion, so its spectrum should be devoid of complicating factors like near-IR excess veiling and accretion excess.  This source is the ideal candidate for direct measurement of two-temperature photosphere due to starspots.

We extend a new spectral forward modeling framework to include two-temperature photosphere models.  We demonstrate the constraining power of full spectrum fitting with panchromatic optical and near-IR echelle spectroscopy.  We show that relatively small starspot filling factors can be detected due to the large spectral grasp of IGRINS.

@@

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{lkca4_lores} 
	\vspace{-100mm}
	\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
	\label{fig:lores}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{tiovar} 
	\vspace{-100mm}
	\caption{Variability in TiO bands measured with ESPaDOnS (see also Table X).  The V-band emission is estimated from fits to the ASAS-SN lightcurve obtained during the same period.  The main panel shows a correlation between V-band magnitude and the TiO-7140 index, while the inset shows a similar correlation with the average of the TiO 6200, CaH 6800, and TiO 7600 indices.}
	\label{fig:tiovar}
\end{figure*}

\section{Testing and runing the mixture model}

Here are some specifics about what we did.

>>>>

\title{Measuring starspot properties by forward modeling IGRINS spectra of \name}

@@

%\subsection{Pre Main Sequence HR diagram spread}

For a single presumably coeval stellar cluster, pre main-sequence HR diagrams exhibit large ($\delta t_{\ast} \sim t_{\ast}$) apparent spreads in age \citep[e.g.][]{2011A&A...534A..83R}. The cause of the large spread in pre main-sequence HR diagrams is controversial: some take it as evidence for intrinsic age spreads, some take it as evidence for non static accretion history \citep{2009ApJ...702L..27B, 2010ARA&A..48..581S}.

@@

%\subsection{Previous studies of Starspots}

Sunspots are probably responsible for biases in stellar effective temperatures derived by different methods.  For example, the APOGEE spectrograph ($1.5-1.70 \;\mu$m at $R=22,500$) measured effective temperatures for 3493 young stars finding offsets in $\teff$ of typically 200$-$500 K and as high as 1000 K compared to previous studies \citep{2014ApJ...794..125C}.  Low resolution optical spectroscopy shows a typical spread of 200 K in the spectral type-to-effective-temperature conversion scale \citep{2014ApJ...786...97H}.  These discrepancies in $\teff$ could be attributable to measurements of spectral features steming from patches of stellar photosphere with different effective temperatures than the bulk photosphere.  Single temperature photospheres are not realistic approximations for complex young photospheres possessing large areal coverage fractions of starspots.

Stellar evolution models including the effect of starspots can make a coeval 10 Myr population exhibit apparent age spreads of 3$-$10 Myr, with derived masses biased towards lower masses \citep{2015ApJ...807..174S}.  The importance of starspots needs to be empirically evaluated to assess systematic biases of pre-main sequence stellar evolutionary models.
%...develop this point more

Modeling the periodic light curves of spotted stars can answer these questions: ``What evidence is there to support the existence of starspots?  What joint constraint can be placed upon the effective temperature contrast and areal coverage of the starspots?  What is the longitudinal asymmetry of the spots?''.  Cyclical light curve variations indicate only the longitudinally asymmetric component of the star spots \citep{2012AJ....143....4H}.  Photometric modeling of the light curve taken alone can only yield degenerate estimates in the temperature contrast and areal coverage fraction.  A star possessing longitudinally symmetric distribution of starspots, or a pole-on star with a single spot, or a non-rotating (or very slowly rotating) star would all show zero or secular amplitude of photometric variability, yet they all contain star spots.

By modeling the composite \emph{spectrum} of a spotted star, it is possible to directly measure the collective areal coverage and effective temperature contrast.  Spectroscopic modeling can still measure the starspot properties in pole-on stars, stars with longitudinally symmetric stars, and non-rotating stars.

@@

%\subsection{Previous studies of \name}

In this paper we characterize the starspot coverage and temperature contrast of \name using full spectrum fitting.  We chose \name for its likelihood of large areal coverage of starspots, as evinced by large-amplitude photometric variability  detected in $BVRI$ bands.  Its period is 3.37$\pm$0.01 days \citep{1993AJ....106.1608V,1994IBVS.4042....1G,grankin08,xiao12}.

Recent optical spectro-polarimetry showed evidence for hot or cool starspots covering an estimated $25\%$ of the stellar surface \citep{2014MNRAS.444.3220D}.  
%This statement needs to be fact-checked, maybe some more supporting info.

The source has little or no accretion ($\log{\dot M} < 8.1$), no detection of mid-infrared nor mm excess, and no stellar companion, so its spectrum should be devoid of complicating factors like near-IR excess veiling and accretion excess.  This source is the ideal candidate for direct measurement of two-temperature photosphere due to starspots.
% These claims need references.

We extend a new spectral forward modeling framework to include two-temperature photosphere models.  We demonstrate the constraining power of full spectrum fitting with panchromatic optical and near-IR echelle spectroscopy.
% Consider putting in Section references here

@@

The details of the methodology are contained in Appendix \ref{methods-details}.  Here we overview the salient concepts needed to follow the results and discussion.

We extended the modular framework for spectral inference described in \iancze \footnote{The open source codebase and its full revision history is available at https://github.com/iancze/Starfish.  The experimental fork discussed in this paper is at https://github.com/gully/Starfish}.  The \iancze technique forward models emulated synthetic spectra from pre-computed model grids, taking into account the discretization error attributable to the coarsely sampled stellar intrinsic parameters.  We use the \PHOENIX~ model grid. We made two principal changes to the modular framework: 1) the introduction of a starspot mixture model characterized by a second $\teff$ and areal fill factor $f$, and 2) the switch from Gibbs sampling to ensemble sampling with chunking by spectral order.  We discuss these two changes below.

Following and extending the notation of \iancze, we can write down the mixture model as:

where $\vM$ is the pixel-level forward model of a spectrum, and we assume the stellar photosphere is characterized as two photospheric components with different effective temperatures, $\teff$ and solid angular coverages $\Omega$, but sharing all the same intrinisic and extrinsic stellar parameters otherwise.  The observed filling factor of photosphere $b$ is simply the ratio:

\begin{eqnarray} \label{eqn:mix_M}
f = \frac{\Omega_b}{\Omega_a + \Omega_b}
\end{eqnarray}

The second important change to the spectral inference framework involves technical aspects of switching from sampling the nuisance and stellar parameters separately in a blocked Gibbs framework to ensemble sampling using \texttt{emcee} \cite{2013PASP..125..306F}.  The affine-invariant ensemble sampler is more resilient to correlations expected among starspot temperature and fill factor than the Metropolis-Hastings sampler used in Gibbs sampling.  The practical effect of this switch is that we obtain unique stellar parameter estimates $\vT_{o}$ for each spectral order $o$, whereas the \iancze strategy had the power to provide a single set of stellar parameters $\vT$ that was based on all $N_{\rm ord}$ spectral orders.  There is no principled way to combine the $N_{\rm ord}$ sets of inferences on $\vT_{o}$, but there are ample heuristics like taking the weighted average of the $o$ sets of marginalized posterior distributions functions.

\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{figures/lkca4_lores} 
	\vspace{-100mm}
	\caption{The low-resolution optical/near-IR spectrum of LkCa 4 is }
	\label{fig:lores}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{figures/tiovar} 
	\vspace{-100mm}
	\caption{Variability in TiO bands measured with ESPaDOnS (see also Table X).  The V-band emission is estimated from fits to the ASAS-SN lightcurve obtained during the same period.  The main panel shows a correlation between V-band magnitude and the TiO-7140 index, while the inset shows a similar correlation with the average of the TiO 6200, CaH 6800, and TiO 7600 indices.}
	\label{fig:tiovar}
\end{figure*}

\section{Methodological details}
\label{methods-details}

@@
